Detection of an Intrinsic Color Parameter within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-09-27 forgerard.tex
plowing on

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
& $X=U$ &$B$&$V$&$R$&$I$\\ \hline
$\alpha_X$
&
$0.0043^{0.0009}_{-0.0009}$
&
$0.0016^{0.0007}_{-0.0007}$
&
$0.0015^{0.0006}_{-0.0006}$
&
$0.0015^{0.0005}_{-0.0005}$
&
$0.0026^{0.0004}_{-0.0005}$
\\
${\alpha_X/\alpha_V-1}$
&
$   1.9^{   1.0}_{  -0.5}$
&
$   0.1^{   0.1}_{  -0.2}$
&
0
&
$  -0.0^{   0.1}_{  -0.1}$
&
$   0.7^{   0.7}_{  -0.3}$
\\
$\beta_X$
&
$ 0.032^{ 0.003}_{-0.003}$
&
$ 0.025^{ 0.003}_{-0.002}$
&
$ 0.025^{ 0.002}_{-0.002}$
&
$ 0.020^{ 0.002}_{-0.002}$
&
$ 0.019^{ 0.002}_{-0.002}$
\\
${\beta_X/\beta_V-1}$
&
$  0.25^{  0.05}_{ -0.05}$
&
$ -0.01^{  0.03}_{ -0.03}$
&
0
&
$ -0.19^{  0.01}_{ -0.01}$
&
$ -0.23^{  0.03}_{ -0.03}$
\\
$\eta_X$
&
$-0.0002^{0.0011}_{-0.0011}$
&
$0.0001^{0.0009}_{-0.0009}$
&
$0.0005^{0.0008}_{-0.0008}$
&
$0.0005^{0.0007}_{-0.0007}$
&
$-0.0003^{0.0006}_{-0.0006}$
\\
${\eta_X/\eta_V-1}$
&
$ -0.38^{  2.21}_{ -1.84}$
&
$ -0.32^{  1.51}_{ -1.15}$
&
0
&
$ -0.11^{  0.28}_{ -0.26}$
&
$ -0.86^{  1.69}_{ -1.24}$
\\
${\gamma_X/\gamma_V-1}$
&
$  0.66^{  0.06}_{ -0.05}$
&
$  0.35^{  0.03}_{ -0.03}$
&
0
&
$ -0.23^{  0.01}_{ -0.01}$
&
$ -0.45^{  0.03}_{ -0.03}$
\\
$R_X$
&
$  4.79^{  0.49}_{ -0.43}$
&
$  3.88^{  0.40}_{ -0.35}$
&
$  2.88^{  0.31}_{ -0.28}$
&
$  2.22^{  0.25}_{ -0.23}$
&
$  1.60^{  0.21}_{ -0.20}$
\\
${\delta_X/\delta_V-1}$
&
$ -0.31^{  0.18}_{ -0.19}$
&
$ -0.17^{  0.10}_{ -0.10}$
&
0
&
$ -0.07^{  0.05}_{ -0.04}$
&
$ -0.17^{  0.10}_{ -0.09}$
\\
$R_{\delta X}$
&
$ -4.06^{  1.97}_{ -4.44}$
&
$ -4.92^{  2.03}_{ -4.95}$
&
$ -5.89^{  2.22}_{ -5.67}$
&
$ -5.48^{  2.04}_{ -5.24}$
&
$ -4.86^{  1.82}_{ -4.66}$
\\
$\sigma_X$
&
$ 0.061^{ 0.013}_{-0.012}$
&
$ 0.034^{ 0.007}_{-0.007}$
&
$ 0.021^{ 0.004}_{-0.006}$
&
$ 0.013^{ 0.007}_{-0.008}$
&
$ 0.044^{ 0.005}_{-0.004}$
\\
\hline
\end{tabular}
\caption{68\% credible Intervals for the Global Fit Parameters \label{global:tab}}
\end{table}

@@

@@

\section{Interpreting $k$ and $D$ As Parameters of a Single Dust Model}
%ipdb> print y, y-ymin, ymax-y
%2.23720862938 0.00612061774367 0.00623996846008
%ipdb> (y,ymin,ymax)  = numpy.percentile(samples[:,1],(50,50-34,50+34)) 
%ipdb> print y, y-ymin, ymax-y
%-3.46707773157e-05 0.000555359008477 0.000616592111953
Until now we have considered $k$ and $D$ as independent external and internal parameters.  However, they may be both
parameters of a single underlying dust.  Indeed the CCM dust law can be exactly written as a two-parameter linear model
\begin{equation}
A_X = A_V a(X) + E(B-V) b(X).
\end{equation}
The Fitzpatrick model can also approximately be described by a linear model in the range of $R_V$ and wavelength under consideration:
the results shown here vary only at the percent level when making this approximation.

In our model, variations in $k$ and $D$ affect magnitudes in two directions in magnitude space
$\frac{\partial A}{\partial E_\delta(B-V)}$, $\frac{\partial A}{\partial E_\gamma(B-V)}$.  In the Fitzpatrick model
variations on $E^F(B-V)$ and $A^F_V=R^F_VE^F(B-V)$ induce magnitude changes in
in directions
$\frac{\partial A}{\partial E^F(B-V)}$, $\frac{\partial A}{\partial A^F_V}$.
The planes in magnitude-space defined by each pair of vectors are nearly parallel, in that
\begin{equation}
\begin{pmatrix}
\frac{\partial A}{\partial E_\delta(B-V)} \\
\frac{\partial A}{\partial E_\gamma(B-V)}
\end{pmatrix}=
\begin{pmatrix}
0.74 & -5.45 \\
1.15 & 2.84
\end{pmatrix} 
\begin{pmatrix}
\frac{\partial A}{\partial R^F_V} \\
\frac{\partial A}{\partial A^F_V}
\end{pmatrix}+
\vec{\epsilon},
\end{equation}
where the residual $\vec{\epsilon}$''s perpendicular to the Fitzpatrick plane
contribute  5\% and 2\% to the total  length of the vector.

@@

\color{red}
A correlation between Hubble residual and host-galaxy mass
was first noted by \citet{2010ApJ...715..743K,2010MNRAS.406..782S}, a signal confirmed to exist in the SNfactory
sample \citep{2013ApJ...770..108C}.
This host-mass bias could be the result of a parameter that was not accounted for in the inference of SN~Ia absolute magnitude.
\citet{2016arXiv160904470M} find that with intrinsic color scatter, the null mass-step effect falls on the 95\%-level of their posterior.
To explore this possibility we plot in Figure~\ref{childress:fig} our intrinsic color excess $E_\gamma(B-V)$  versus host mass 
\citep{2016rigault}.  Choosing the same mass cut as  \citet{2013ApJ...770..108C}, $\log{(M/M_\sun)}=10$, we calculate the average
color excess for low- (high)-mass hosts to be
$  -0.0043 \pm    0.0022$
($  -0.0072 \pm    0.0073$).  
The ideograms for the low- and high-mass samples is shown in Figure~\ref{childress:fig}.
The 2-sample Kolmogorov-Smirnov test of the medians gives a $p$-value of
%---
0.29.
%---
From the above evidence, we cannot show that low- and high-mass hosts produce SNe~Ia with difference $E_\gamma(B-V)$ distributions. 
Using the SNfactory sample,
\citet{2013A&A...560A..66R} find that step in Hubble residuals as better related to local star formation rate, rather than host mass.
The connection between our intrinsic parameter and local star formation and other host properties is deferred to
\citet{2016rigault}.
\color{black}
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/rigault3.pdf}
      \includegraphics[width=2.8in]{output11/rigault2.pdf}
      \caption{\textcolor{red}{Left: Intrinsic parameter $E_\gamma(B-V)$  versus host mass.  Hosts too faint for a mass measurement
      are assigned a mass of 5.
Right: Ideograms of  $E_\gamma(B-V)$ for supernovae in $\log{(M/M_\sun)}<10$ and $\log{(M/M_\sun)}>10$ hosts. 
}
   \label{childress:fig}}
\end{figure}

>>>>

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
& $X=U$ &$B$&$V$&$R$&$I$\\ \hline
$\alpha_X$
&
$0.0043^{0.0009}_{-0.0009}$
&
$0.0016^{0.0007}_{-0.0007}$
&
$0.0015^{0.0006}_{-0.0006}$
&
$0.0015^{0.0005}_{-0.0005}$
&
$0.0026^{0.0005}_{-0.0004}$
\\
${\alpha_X/\alpha_2-1}$
&
$   1.9^{   1.0}_{  -0.5}$
&
$   0.1^{   0.1}_{  -0.2}$
&
$   0.0^{   0.0}_{   0.0}$
&
$  -0.0^{   0.1}_{  -0.1}$
&
$   0.7^{   0.7}_{  -0.3}$
\\
$\beta_X$
&
$ 0.032^{ 0.003}_{-0.003}$
&
$ 0.025^{ 0.002}_{-0.003}$
&
$ 0.025^{ 0.002}_{-0.002}$
&
$ 0.020^{ 0.002}_{-0.002}$
&
$ 0.019^{ 0.002}_{-0.002}$
\\
${\beta_X/\beta_2-1}$
&
$  0.26^{  0.05}_{ -0.05}$
&
$ -0.01^{  0.03}_{ -0.03}$
&
$  0.00^{  0.00}_{  0.00}$
&
$ -0.19^{  0.01}_{ -0.01}$
&
$ -0.24^{  0.03}_{ -0.03}$
\\
$\eta_X$
&
$-0.0002^{0.0012}_{-0.0011}$
&
$-0.0000^{0.0010}_{-0.0009}$
&
$0.0005^{0.0008}_{-0.0008}$
&
$0.0005^{0.0007}_{-0.0007}$
&
$-0.0003^{0.0006}_{-0.0006}$
\\
${\eta_X/\eta_2-1}$
&
$ -0.40^{  2.33}_{ -1.90}$
&
$ -0.34^{  1.60}_{ -1.20}$
&
$  0.00^{  0.00}_{  0.00}$
&
$ -0.11^{  0.29}_{ -0.27}$
&
$ -0.84^{  1.66}_{ -1.27}$
\\
${\gamma^0_X/\gamma^0_2-1}$
&
$  0.66^{  0.06}_{ -0.05}$
&
$  0.35^{  0.03}_{ -0.03}$
&
$  0.00^{  0.00}_{  0.00}$
&
$ -0.23^{  0.01}_{ -0.01}$
&
$ -0.45^{  0.03}_{ -0.03}$
\\
$R^_{\gamma^0 X}$
&
$  4.78^{  0.47}_{ -0.42}$
&
$  3.87^{  0.38}_{ -0.34}$
&
$  2.87^{  0.30}_{ -0.27}$
&
$  2.21^{  0.25}_{ -0.23}$
&
$  1.58^{  0.21}_{ -0.19}$
\\
${\gamma^1_X/\gamma^1_2-1}$
&
$ -0.31^{  0.17}_{ -0.18}$
&
$ -0.17^{  0.09}_{ -0.10}$
&
$  0.00^{  0.00}_{  0.00}$
&
$ -0.07^{  0.05}_{ -0.04}$
&
$ -0.17^{  0.10}_{ -0.09}$
\\
$R_{\gamma^1 X}$
&
$ -4.02^{  1.87}_{ -4.10}$
&
$ -4.87^{  1.93}_{ -4.54}$
&
$ -5.84^{  2.11}_{ -5.21}$
&
$ -5.45^{  1.94}_{ -4.83}$
&
$ -4.86^{  1.75}_{ -4.32}$
\\
$\sigma_X$
&
$ 0.060^{ 0.012}_{-0.012}$
&
$ 0.033^{ 0.007}_{-0.007}$
&
$ 0.021^{ 0.004}_{-0.006}$
&
$ 0.012^{ 0.007}_{-0.008}$
&
$ 0.044^{ 0.005}_{-0.004}$
\\
\hline
\end{tabular}
\caption{68\% credible Intervals for the Global Fit Parameters \label{global:tab}}
\end{table}

@@

\section{Supernova Model Part 2}
\label{model2:sec}
We hypothesize that at peak brightness
SN~Ia broadband magnitudes and colors are correlated with
spectral features: equivalent widths and line positions are considered as statistics localized in wavelength that are insensitive to variations in
broadband colors.
Spectra may not efficiently capture all variations in broadband magnitude, so an additional source
of intrinsic variation is accommodated.
The intrinsic spectral and color parameters may not deterministically predict magnitudes, but rather do so with some intrinsic dispersion.
\textcolor{red}{The intrinsic magnitudes are then
modified by an extrinsic physical process (i.e.\ dust) to produce apparent magnitudes.}

We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
\textcolor{red}{ and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature}:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
$EW_{Si}$ at peak brightness is an effective proxy for the light-curve shape parameter
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
An intrinsic-color parameter $D$ describes  intrinsic magnitude/color variation unaccounted for by the spectral features.
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset $\Delta$ is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on an
extrinsic-color parameter $k$.  The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $v_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
The model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} + \delta_0 D\\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si} + \delta_1 D \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} + \delta_2 D\\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} + \delta_3 D\\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si} + \delta_4 D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$, $\eta$, and $\delta$  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters and intrinsic-color parameter.
(In this article we freely interchange the parameter indices with  $01234$ or $UBVRI$.)
The global parameters $\gamma$ are the slopes that connect the extrinsic-color
parameter and observed magnitudes.

To constrain the degrees of freedom and degeneracies inherent in the model we impose that
\color{red}
each of the vectors $k+1/N$ and $D+1/N$ are simplexes, where $N$ is the number of objects in our sample: these
conditions break the 
degeneracies  $\gamma \rightarrow a\gamma$, $k \rightarrow a^{-1}k$ and $\delta \rightarrow a\delta$, $D \rightarrow a^{-1}D$.
Furthermore we impose
\color{black} 
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0, \langle D \rangle=0, \gamma^0_{0} > 0, \gamma^1_{0} < 0, \delta \cdot \gamma^0=0, \delta \cdot \gamma^1=0.
\end{equation}
The first two specify the definition of zero color relative to which the color excess is measured.    The latter two exclude degenerate posterior space
associated
with  simultaneous sign flips of
$\delta$--$D$ and $\gamma$--$k$
\color{red}
(when running our fits, $k$ and $D$ are set to
identical initial conditions, so the difference in the signs of $\gamma_0$ and $\delta_0$ distinguishes the two.)
Although not formally a degeneracy in the model, practically the parameter pairs  $\delta$--$D$ and $\gamma$--$k$ are degenerate with
each other, as is apparent after 
running the Monte Carlo with exchange of the initial  and boundary conditions of $\delta$ and $\gamma$.
The different initial and boundary conditions for $\gamma_0$ and $\delta_0$ break the degeneracy of our mixture model;
otherwise identical initial conditions for $\gamma$, $\delta$ produce indistinguishable posteriors
with relatively broad credible intervals and lower maximum likelihood.
\color{black}
As the model cannot distinguish which effect is intrinsic or extrinsic, we do so
by assigning the parameters whose values are consistent with dust to $\gamma$ and the others to $\delta$.

@@

\section{Interpreting $k$ and $D$ As Parameters of a Single Dust Model}
%ipdb> print y, y-ymin, ymax-y
%2.23720862938 0.00612061774367 0.00623996846008
%ipdb> (y,ymin,ymax)  = numpy.percentile(samples[:,1],(50,50-34,50+34)) 
%ipdb> print y, y-ymin, ymax-y
%-3.46707773157e-05 0.000555359008477 0.000616592111953
Until now we have considered $k$ and $D$ as independent external and internal parameters.  However, they may be both
parameters of a single underlying dust.  Indeed the CCM dust law can be exactly written as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The Fitzpatrick model can also approximately be described by a linear model in the range of $R_V$ and wavelength under consideration:
the results shown here vary only at the percent level when making this approximation.
Similarly our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2}  E_{\gamma^1}(B-V).
\end{equation}

In our model, variations in $k$ and $D$ affect magnitudes in two directions in magnitude space
$\frac{\partial A}{\partial E_\delta(B-V)}$, $\frac{\partial A}{\partial E_\gamma(B-V)}$.  In the Fitzpatrick model
variations on $E^F(B-V)$ and $A^F_V=R^F_VE^F(B-V)$ induce magnitude changes in
in directions
$\frac{\partial A}{\partial E^F(B-V)}$, $\frac{\partial A}{\partial A^F_V}$.
The planes in magnitude-space defined by each pair of vectors are nearly parallel, in that
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
1.16 & 2.82 \\
0.74 & -5.43
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\vec{\epsilon},
\end{equation}
where the residual $\vec{\epsilon}$''s perpendicular to the Fitzpatrick plane
contribute  2\% and 5\% to the total  length of the vector.

@@

\color{red}
A correlation between Hubble residual and host-galaxy mass
was first noted by \citet{2010ApJ...715..743K,2010MNRAS.406..782S}, a signal confirmed to exist in the SNfactory
sample \citep{2013ApJ...770..108C}.
This host-mass bias could be the result of a parameter that was not accounted for in the inference of SN~Ia absolute magnitude.
\citet{2016arXiv160904470M} find that with intrinsic color scatter, the null mass-step effect falls on the 95\%-level of their posterior.
To explore this possibility we plot in Figure~\ref{childress:fig} our intrinsic color excess $E_\gamma(B-V)$  versus host mass 
\citep{2016rigault}.  Choosing the same mass cut as  \citet{2013ApJ...770..108C}, $\log{(M/M_\sun)}=10$, we calculate the average
color excess for low- (high)-mass hosts to be
$  -0.0043 \pm    0.0022$
($  -0.0072 \pm    0.0073$).  
The ideograms for the low- and high-mass samples is shown in Figure~\ref{childress:fig}.
The 2-sample Kolmogorov-Smirnov test of the medians gives a $p$-value of
%---
0.29.
%---
From the above evidence, we cannot show that low- and high-mass hosts produce SNe~Ia with difference $E_\gamma(B-V)$ distributions. 
Using the SNfactory sample,
\citet{2013A&A...560A..66R} find that step in Hubble residuals as better related to local star formation rate, rather than host mass.
The connection between our intrinsic parameter and local star formation and other host properties is deferred to
\citet{2016rigault}.
\color{black}
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
%   \includegraphics[width=2.8in]{output11/rigault3.pdf}
%      \includegraphics[width=2.8in]{output11/rigault2.pdf}
      \caption{\textcolor{red}{Left: Intrinsic parameter $E_\gamma(B-V)$  versus host mass.  Hosts too faint for a mass measurement
      are assigned a mass of 5.
Right: Ideograms of  $E_\gamma(B-V)$ for supernovae in $\log{(M/M_\sun)}<10$ and $\log{(M/M_\sun)}>10$ hosts. 
}
   \label{childress:fig}}
\end{figure}
